---
title: "Gamma Distribution"
author: "Elena Sophia"
date: "January 22, 2017"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

suppressMessages(library(shiny))
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
rm(list = ls())
```
##

Gamma distribution is a two-parameter family of continuous probability distributions. The common exponential distribution and chi-squared distribution are special cases of the gamma distribution.

Parameters:   

- $k$ > 0   shape
- $theta$ > 0   scale

PDF:                ${\frac {1}{\Gamma (k)\theta ^{k}}}x^{k\,-\,1}e^{-{\frac {x}{\theta }}}$



Mean:             $\scriptstyle \mathbf {E} [X]=k\theta$

Variance:         $\scriptstyle\operatorname{Var}[ X] = k\theta^2$




Source: Wikipedia

##  Gamma Distribution Applications

- model errors in multi-level Poisson regression models, because the combination of the Poisson distribution and a gamma distribution is a negative binomial distribution.

 - model the multi-path fading of signal power in wireless communication

- describe the distribution of inter-spike intervals in neuroscience

- model aggregate insurance claims and the amount of rainfall accumulated in a reservoir

- in bacterial gene expression, the copy number of a constitutively expressed protein often follows the gamma distribution.

- in genomics, the gamma distribution was applied in peak calling step in ChIP-chip and ChIP-seq data analysis.



Source: Wikipedia


## 

```{r, echo = FALSE}

  noobs <- c(500)
  Gshape <- c(2)
  Gscale <- c(1)

  # generate sample of random gamma variables

  Gdata <- data.frame(rgamma(noobs,
                             shape = Gshape,
                             scale = Gscale))
  names(Gdata) <- c("Gdata")
  
 #   sample statistics
 samplestats <- Gdata %>%
    summarize(samplemean = round(mean(Gdata), 2),
               samplestdev = round(sd(Gdata), 2),
               samplevar = round(var(Gdata), 2),
               expmean = Gshape * Gscale,
               expstddev = round(sqrt(Gshape * Gscale ^ 2), 2),
               expvar = Gshape * Gscale ^ 2)

   Gmode <- ifelse(Gshape >= 1, (Gshape - 1) * Gscale, Gshape * Gscale)
   Gtop <- dgamma(Gmode, shape = Gshape, scale = Gscale)

   #   Create exponential plot
   gcolors <- c("Gamma Distribution Probability Density" = "red",
                "Gamma Distribution Mean" = "blue",
                "Sample Probability Density" = "black",
                "Sample Mean" = "green")

   ggplot (Gdata, aes ( x = Gdata)) +
     geom_histogram(aes (y = ..density..), fill = "grey", binwidth = .1,
                    show.legend = FALSE) +
     geom_density(aes(color = "Sample Probability Density"),
                  size = 1, show.legend = FALSE) +
     stat_function(fun = dgamma, args = list(shape = Gshape, scale = Gscale),
                   aes(color = "Gamma Distribution Probability Density"),
                   show.legend=TRUE, size = 1) +
     geom_segment(aes(x = Gshape * Gscale, xend = Gshape * Gscale,
                      y = 0, yend = Gtop,
                      color="Gamma Distribution Mean"),
                  show.legend = FALSE, size=1, alpha = .5) +
     geom_segment(aes(x = samplemean, xend = samplemean, y = 0, yend = Gtop,
                      color="Sample Mean"), data = samplestats,
                  show.legend = FALSE, size = 2, linetype = 2, alpha = .8) +
     scale_fill_manual(name = "", values = gcolors) +
     scale_colour_manual(name="",values=gcolors) +
     theme_bw() + geom_rug() +
     labs (title = substitute(paste("Gamma Distribution Simulation with k = ",
                                    k, ", ", theta, " = ", t),
                              list(k = Gshape, t = Gscale)), x = "x", size = 25) +
     scale_x_continuous(expand = c(0, 0), breaks = seq(0, 35, 5)) +
     theme(plot.title = element_text(lineheight=1, face="bold", vjust = 1, size = 18),
           axis.title = element_text(face="bold", size=15),
           axis.text = element_text(face="bold", size = 12),
           legend.position = c(.75, .85))


  
  
  
```

## 

The plot on the previous slide shows theoretical gamma distribution
with shape $k$ = 2 and scale $\theta$ = 1 and corresponding  $\mu$ = `r samplestats$expmean` and $\sigma$ = `r samplestats$expstddev`.
For 500 iid random gamma variables sample distribution mean $\bar{X}$ = `r samplestats$samplemean` and standard error S = `r samplestats$samplestdev` are close to the underlying population parameters.  The sample density (black line on the plot) follows that of the gamma distribution (shown in red); theoretical and sample means are shown as blue and dashed green line.

